---
title: "AZV methods"
output: html_document
date: "2025-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggdag)
library(tidyverse)
library(dagitty)
library(patchwork)   
```

## R Markdown

```{r Observable}

azv_dag <- dagify(
    V ~  OC + OH + AS + PB,
    AE ~ OC + OH + AS + V,
    OC ~ OH + AS + PB,
    OH ~ OC + AS,
    PB ~ OH + AS + OC,
  exposure = "V",
  outcome = "AE",
  latent = c("PB"),
  labels = c(
             V = "vaccination", 
             OC = "COVID infection",
             OH = "health",
             PB = "pandemic behavior",
             AS = "age & sex", 
             AE = "adverse effect"),
   coords = list(  # Define coordinates inside dagify
    x = c(V = 0, AE = 3,  OC = 2, OH = 1, AS = 1.5, PB = 1.5),
    y = c(V = 1.4, AE = 1.4,  OC = 2, OH = 2, AS = 0.8, PB = 0)
  )
)

p <- ggdag_status(azv_dag, use_labels = "label", text = FALSE) +
  guides(fill = "none", color = "none") +  # Disable the legend
  theme_dag()
ggsave( "dag.pdf", p)
p
print(impliedConditionalIndependencies(azv_dag))
ggdag_paths(azv_dag) + theme_dag()

ggdag_adjustment_set(azv_dag) +theme_dag()
```

azv_dag <- dagify(
    V ~ HC + HH +  OC + OH + AS + PB,
    AE ~ HC + HH + OC + OH + AS + V,
    HC ~ HH + OH + AS + PB,
    OC ~ HH + OH + AS + PB,
    HH ~ HC + OC + AS,
    OH ~ HC + OC + AS,
  exposure = "V",
  outcome = "AE",
  latent = c("HH","HC","PB"),
  labels = c(
             V = "vaccination", 
             HC = "unbserved COVID",
             OC = "observed COVID",
             HH = "unbserved health",
             OH = "observed health",
             PB = "pandemic behavior",
             AS = "age & sex", 
             AE = "adverse effect"),
   coords = list(  # Define coordinates inside dagify
    x = c(V = 0, AE = 3, HC = 1, OC = 2, HH = 2.3, OH = 1, AS = 1.5, PB = 0.3),
    y = c(V = 1.4, AE = 1.4, HC = 0, OC = 2, HH = 0, OH = 2, AS = 0.8, PB = 0)
  )
)
